What's in a name? On the automated topic naming of software maintenance activities

### 2010-03-06 icpc/icpc2010.tex
supervised labelling
changes to intro and refs

  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.  While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.  In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.  We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.  In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as WordNet.  In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS.  Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects. Within the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time. Our automated techniques can label topics without much computational overhead.

@@

 Few would argue that software development is a bed of roses.  To most of us, the important properties of a rose are sensual in nature, and concern appearance, odour, and feel.  While horticulturalists and botanists may see more semantic depth than this, one rose is often just as pleasing to us as another.  Software development artifacts, on the other hand, are abstract and intangible.  They are often unnamed and ephemeral --- even more so than roses --- and maybe partially derived from other unnamed and ephemeral artifacts.  Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our work topics and be able to categorize the development artifacts as belonging to one or more of these topics.

%Original is here:

@@

%To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above. We can then categorize development artifacts as belonging to one or more of these topics.

% This correlates issues with topics
As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

%Topics arise from the wide variety of issues which occurs during a software project's lifecycle. These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

@@

A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system. Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case, so we have decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.

@@

% Clean this up
Traditionally topic extraction has required manual annotation to derive domain-relevant labels. This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy. This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics. We also compare this approach to machine learners.
%We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology. From the annotated topics, we %apply automated classifiers to find labels for the remaining topics.

@@

\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when a developer commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (an \emph{A} is $0.9$, a \emph{C} is $0.6$), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

\subsection{Topic and Concept Extraction}
Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining software repositories, in particular analyzing source code, bugs or developer comments.
\begin{comment}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comment}

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to 1 or more documents. Our problem is that these topics are not easy to interpret and we feel that automatic labelling or naming of these topics would be helpful with respect to interpreting what a topic is about. LDA's topics extracted in an unsupervised manner, LDA relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

@@

One issue that arises with the use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. For example, given the word list ``listener change remove add fire'', Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require some expert in the field to determine them. Our technique is automated, because we match keywords from WordNet to words in the topic model.

\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, supervised learners have the advantage of domain knowledge. This typically produces improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert
In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test various supervised learners on our annotated data-set.

@@

To evaluate our approach, we sought candidate systems that were
mature projects and had openly accessible source control repositories.
We also decided to select systems from the same application
domain, as we felt the functional requirements would probably be
broadly similar.
%We wanted to compare two mature software projects, that had open %repositories, within the same application domain. 
% This allows us to partially control for differences in functional %requirements. 
We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}.
%How many developers?
%There are approximately ... developers working on it. 
The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.

\subsection{Generating the data}
From each project, we used source control commit comments, these are the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.

From that data-set, we created an XML file which windowed commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied an Blei's LDA implementation against the word distributions of these commits, and generated list of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.

@@

Topics are word distributions: essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish and understand.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that a word from our word list. We used several different word lists for comparison.

Our first word list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}. That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. The labels we used:
\begin{quotation}
\small \noindent \textsf{
integrity, security,
interoperability, testability, maintainability, traceability,
accuracy, modifiability, understandability, availability, modularity,
usability, correctness, performance, verifiability, efficiency,
portability, flexibility, reliability.
}
\end{quotation}

Our second word list set, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
%There is some debate about the significance and importance of the %terms in this model. 
%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p. %58]{Bøegh2008},'' which is sufficient for the purposes of this %research. 
ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
However, the terms extracted from ISO9126 may not capture all words associated with the labels.  For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. Therefore we took the words from the taxonomy and expanded them.

@@

Mei et al.~\cite{Mei2007} use context information to automatically name topics. They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. They do not use external domain-specific information as we do.
In \cite{ernst10refsq}, we describe an application similar to this to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.

@@

Most of this related research examined project source code. Software mining of other project corpora is less common. Similarities exist with approaches that begin with structured taxonomies, as with the Hismo software ontology~\cite{girba06}. Rigby and Hassan~\cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.

@@

In conclusion we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics. Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet. We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.

>>>>

  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.  While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.  In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.  We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.  In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as WordNet.  In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS.  Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects. With   in the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time. Our automated techniques can label topics without much computational overhead.

@@

Few would argue that software development is a bed of roses.  To most of us, the important properties of a rose are sensual in nature, and concern appearance, odour, and feel.  While horticulturalists and botanists may see more semantic depth than this, one rose is often just as pleasing to us as another.  Software development artifacts, on the other hand, are abstract and intangible.  They are often unnamed and ephemeral --- even more so than roses --- and maybe partially derived from other unnamed and ephemeral artifacts.  Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our work topics and be able to categorize the development artifacts as belonging to one or more of these topics. 
%Original is here:

@@

% To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above. We can then categorize development artifacts as belonging to one or more of these topics.

Topics arise from the wide variety of issues which occur during a software project's lifecycle. These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

@@

A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system. Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case. We have therefore decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.

@@

% Clean this up
Traditionally topic extraction has required manual annotation to derive domain-relevant labelling information. This paper attempts to implement \emph{labelled topic extraction}, in which topic are extracted and given labels relating to a cross-project taxonomy. This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms in our topic. We also compare this approach to machine learners.
%We start with an unsupervised topic learning algorithm, then manually %annotate the topics using the ontology. From the annotated topics, we %apply automated classifiers to find labels for the remaining topics.

@@

\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top ten most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

\subsection{Topic and Concept Extraction}
Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
\begin{comment}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comment}

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents. Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear. We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

@@

One issue that arises with use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, as well as in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. Our technique is automated, because we match keywords from WordNet~\cite{Fellbaum1998} to words in the topic model.

\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices. In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.

@@

To evaluate our approach, we sought candidate systems that were
mature projects and had openly accessible artifact repositories.
We also decided to select systems from the same application
domain, as we felt the functional requirements would probably be
broadly similar.
%We wanted to compare two mature software projects, that had open %repositories, within the same application domain. 
% This allows us to partially control for differences in functional %requirements. 
We selected MySQL and MaxDB as they were open-source, partially-commercial database systems. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}.
%How many developers?
%There are approximately ... developers working on it. 
The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.

\subsection{Generating the data}
From each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as common English words \emph{the} and \emph{at}.

From that data-set, we created an XML file which separated commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated list of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.

@@

Topics are word distributions: essentially lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words. We used several different word lists for comparison.

Our first list, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}. That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. The labels we used:
\begin{quotation}
\small \noindent \textsf{
integrity, security,
interoperability, testability, maintainability, traceability,
accuracy, modifiability, understandability, availability, modularity,
usability, correctness, performance, verifiability, efficiency,
portability, flexibility, reliability.
}
\end{quotation}

Our second list, \textsf{exp2}, relied on the ISO quality model (ISO9126)~\cite{iso9126}. ISO9126 describes six high-level quality requirements (listed in Table \ref{tbl:wnsig}). 
%There is some debate about the significance and importance of the %terms in this model. 
%However, it is ``an international standard and thus provides an %internationally accepted terminology for software quality \cite[p. %58]{Bøegh2008},'' which is sufficient for the purposes of this %research. 
ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Bøegh2008},'' that is sufficient for the purposes of this research.
However, the terms extracted from ISO9126 may not capture all words associated with the labels.  For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. Therefore we took the words from the taxonomy and expanded them.

@@

Mei et al.~\cite{Mei2007} use context information to automatically name topics. They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. They do not use external domain-specific information as we do.
In \cite{ernst10refsq}, we describe our earlier project, similar to this, to identify change in quality requirements in GNOME software projects; our approach is solely text-matching, however, and does not leverage machine learning strategies.

@@

% Most of this related research examined project source code. Software mining of other project corpora is less common. Similarities exist with approaches that begin with structured taxonomies, as with the Hismo software ontology~\cite{girba06}. Rigby and Hassan~\cite{rigby07msr} used a general purpose taxonomy to classify developer email according to temperament.

@@

We have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics. Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet. We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.
